Current and future constraints on primordial magnetic fields

### 2016-10-25 pmf.tex
a

We present new limits on the amplitude of potential primordial magnetic fields (PMFs) using temperature and polarization measurements of the cosmic microwave background (CMB)  from \planck{}, \bicepkeck{}, \pb, and \sptpol. 
We reduce by a factor of 2.5 the upper limit on the CMB anisotropy power due to a PMF, from $\apmf < 0.73$ from Planck alone to $\apmf < 0.30$ for the combined dataset at 95\% CL. 
%Most of the improvement is due to the addition of the \bicepkeck{} data; without these bandpowers, the combined limit weakens to $\apmf < 0.\tbd{XX}$. 
We also forecast the expected future limits from stage III CMB experiments (like \sptnew{},  \advactpol, or the \simons) and the proposed CMB stage IV experiment. 
Future CMB experiments should dramatically reduce the current uncertainties, by two orders of magnitude for stage III experiments and three orders of magnitude for stage IV. 
The constraints from a stage IV experiment have the potential to rule out much of the parameter space for PMFs.

@@

@@

The experiments that will begin taking data next year will dramatically improve constraints on the PMF power. 
With the minimal cosmological model, the 1-sigma forecasts for \planck{}+\ho{} is $\sigma(\apmf)=0.47$, slightly worse than indicated by the upper limits found with the full \planck{} likelihood. 
The uncertainty falls nearly 400-fold by adding the two stage III experiments in this simplified model. 
Including all eleven parameters marginally weakens the \planck+\ho{} constraints on \apmf{} by 23\%. 
Parameter degeneracies degrade the Stage III constraint by a substantial factor (x2.5). 
Still in this 11-parameter model, the addition of stage III CMB experiments improves the \apmf{} uncertainty by a factor of $\sim$180 to  $\sigma(\apmf)=0.0031$. 
We can expect substantially tighter constraints on PMFs by the end of the decade.

@@

The primary motivation behind the proposed CMB stage IV experiment is to search for inflationary gravitational waves and neutrino masses, however, the experiment should also enable extremely sensitive searches for PMFs.

 extremely powerful

>>>>

We present new limits on the amplitude of potential primordial magnetic fields (PMFs) using temperature and polarization measurements of the cosmic microwave background (CMB)  from \planck{}, \bicepkeck{}, \pb, and \sptpol. 
We reduce twofold the upper limit on the CMB anisotropy power due to a PMF, from $\apmf < 0.68$ from Planck alone to $\apmf < 0.33$ for the combined dataset at 95\% CL. 
%Most of the improvement is due to the addition of the \bicepkeck{} data; without these bandpowers, the combined limit weakens to $\apmf < 0.\tbd{XX}$. 
We also forecast the expected future limits from stage III CMB experiments (like \sptnew{},  \advactpol, or the \simons) and the proposed CMB stage IV experiment. 
Future CMB experiments should dramatically reduce the current uncertainties, by two orders of magnitude for stage III experiments and three orders of magnitude for stage IV. 
The constraints from a stage IV experiment have the potential to rule out much of the parameter space for PMFs.

@@

As a sanity check, we look at the Fisher matrix forecasts for Planck alone as well in both of these cases. 
For both chains,  we look at the sigma in MCMC with \apmf{} allowed to go negative. 
While a negative \apmf{} is unphysical, this insulates the analysis against the data preferring negative values, and thereby producing tighter limits. 
We find for \lcdm{}  + \apmf{}: $\sigma(\apmf) = XX$ (to be compared to YY from the Fisher matrix). 
We find for the 11-parameter model: $\sigma(\apmf) = XX$ (to be compared to YY from the Fisher matrix). 
This degree of consistency is reasonable. \tbd{}

@@

The experiments that will begin taking data next year will dramatically improve constraints on the PMF power. 
With the minimal cosmological model, the 1-sigma forecasts for \planck{}+\ho{} is $\sigma(\apmf)=0.47$ as noted before.  
The uncertainty is forecast to fall nearly 400-fold by adding the two stage III experiments in this simplified model to $\sigma(\apmf)=0.0012$. 
Including all eleven parameters marginally weakens the \planck+\ho{} constraints on \apmf{} by 23\%, and the stage III constraint by a substantial factor of 2.5. 
Even so within this 11-parameter model, the addition of stage III CMB experiments improves the \apmf{} uncertainty by a factor of $\sim$\,180 to  $\sigma(\apmf)=0.0031$. 
We can expect substantially tighter constraints on PMFs by the end of the decade.

@@

The primary motivation behind the proposed CMB stage IV experiment is to search for inflationary gravitational waves and neutrino masses, however, the experiment should also enable extremely sensitive searches for PMFs. 
We begin by making forecasts for the fiducial CMB-S4 configuration as laid out in Table~\ref{tab:experiments}.
